# Quantitative Results

```{r, include=FALSE}
library(here)
library(tidyverse)
library(readxl)
library(jtools)
library(cowplot)
library(kableExtra)
library(knitr)
library(huxtable)
```

This chapter summarizes the results of the factor analysis, index development, 
index validation, and the development of a combined index to describe site-level
variation in urban quality.

## Factor analysis

Both the Hull method and the Kaiser-Guttman
criterion suggested a five-factor solution would
be appropriate for out data set. Figure \@ref(fig:hull-figure) illustrates 
the results of the Hull method with a plot of 
goodness-of-fit versus degrees of freedom for potential solutions with numbers 
of factors ranging from zero to fourteen. Figure \@ref(fig:kgc-figure) illustrates that there are 
five eigenvalues greater than one, suggesting a 
five-factor solution according to the Kaiser-Guttman Criterion.

```{r hull-figure, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Results of Kaiser-Guttman criterion for determining the number of factors'}
here("04_figures",
     "hull-n-factors.png") %>%
knitr::include_graphics()
```

```{r kgc-figure, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Results of Hull method for determining the number of factors'}
here("04_figures",
     "KGC-n-factors.png") %>%
knitr::include_graphics()
```

We assigned names to each factor based on a visual 
inspection of the results. The _drivable_ factor had the highest loadings
for variables representing access by car to most destination types. The
_walkable_ factor has high loadings for variables representing access
by walking and transit. The _diverse_ index is characterized by diversity of 
people (high percentages of black residents and low percentages of white
residents), diversity of land use (a greater number of distinct land
uses in the immediate vicinity and a shorter average distance to disamenities), 
and lower assessed property values. The _dense_ factor is characterized by lower
values for the radius of the smallest buffer containing two thousand residents 
(i.e. higher population densities) and higher access to retail and grocery 
locations by non-motorized modes. The _amenities_ factor is characterized 
by non-motorized and transit access to retail and grocery locations. Figure
\@ref(fig:loading-fig) illustrates the loadings of each individual variable 
onto each of the five factors.

```{r loading-fig, echo = FALSE, fig.asp=1.9, out.width='100%', fig.cap='Factor loadings'}
# library(extrafont)
# 
# font_import()
# font_install("fontcm")
# loadfonts()

loadings <- here("02_data",
                 "loadings.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(order = case_when(abs(f_drivable) > 0.2 ~ abs(f_drivable) * 10^5,
                           abs(f_walkable) > 0.2 ~ abs(f_walkable) * 10^4,
                           abs(f_diverse) > 0.2 ~ abs(f_diverse) * 10^3,
                           abs(f_dense) > 0.2 ~ abs(f_dense) * 10^2,
                           abs(f_amenities) > 0.2 ~ abs(f_amenities) * 10,
                           TRUE ~ f_amenities)) 

long_loadings <- loadings  %>%
  pivot_longer(c(-variable, -order), 
               names_to = "Factor", 
               values_to = "Loading") %>%
  mutate(primary = abs(Loading) > 0.4) %>%
  mutate(Factor = factor(Factor, 
                         levels = c("f_drivable",
                                    "f_walkable",
                                    "f_diverse",
                                    "f_dense",
                                    "f_amenities"),
                         labels = c("Accessibility by\ndriving",
                                    "Accessibility by\nwalking",
                                    "Diversity of\npeople and\nbuildings",
                                    "Accessibility by\ncycling",
                                    "Accessibility to\nshopping \n(food and retail)"))) 

figure_labels <- rev(c("Car access to hospitality jobs",
                              "Car acces to high-paying jobs",
                              "Car access to all jobs",
                              "Car access to low-paying jobs",
                              "Car access to entertainment jobs",
                              "Car access to destination parcels",
                              "Car access to schools",
                              "Car access to parks",
                              "Car access to retail jobs",
                              "Bike access to high-paying jobs",
                              "Car access to grocery stores",
                              "Bike access to all jobs",
                              "Bike access to hospitality jobs",
                              "Bike access to low-paying jobs",
                              "Bike access to destination parcels",
                              "Walking access to destination\nparcels",
                              "Hispanic percentage of nearest\n2,000 neighbors",
                              "Bike access to entertainment jobs",
                              "Transit access to destination\nparcels",
                              "Walking access to high-paying jobs",
                              "Walking access to all jobs",
                              "Transit access to all jobs",
                              "Transit access to high-paying jobs",
                              "Transit access to low-paying jobs",
                              "Walking access to low-paying jobs",
                              "Walking access to hospitality jobs",
                              "Transit access to hospitality jobs",
                              "Walking access to entertainment\njobs",
                              "Transit access to entertainment\njobs",
                              "Transit access to retail jobs",
                              "Walking access to retail jobs",
                              "Land value (log-transformed)",
                              "Property value (log-transformed)",
                              "Black percentage of nearest\n2,000 neighbors",
                              "Non-Hispanic white percentage of\nnearest 2,000 neighbors",
                              "Number of land uses in 20,000-\nneighbor buffer",
                              "Average distance to ten disamenities\n(log-transformed)",
                              "Lot area (log-transformed)",
                              "Walking access to parks",
                              "Transit access to parks",
                              "Bike access to retail jobs",
                              "Recent sale price (inflation-adjusted\nand log-transformed)",
                              "Bike access to schools",
                              "Bike access to parks",
                              "Bike access to grocery stores",
                              "Walking access to schools",
                              "Radius of buffer containing 2,000\nneighbors",
                              "Transit access to schools",
                              "Walking access to grocery stores",
                              "Transit access to grocery stores"))

plot <- ggplot(long_loadings) +
  geom_point(aes(x = Loading,
                 y = reorder(variable, order),
                 color = Factor,
                 alpha = abs(Loading))) +
  facet_wrap(facet = vars(Factor), ncol = 5) +
  scale_y_discrete(name = "",
                   labels = figure_labels) +
  scale_x_continuous(name = "Variable loading") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

plot

big_plot <- ggplot(long_loadings) +
  geom_point(size = 5,
             aes(x = Loading,
                 y = reorder(variable, order),
                 color = Factor,
                 alpha = abs(Loading))) +
  facet_wrap(facet = vars(Factor), ncol = 5) +
  scale_y_discrete(name = "",
                   labels = figure_labels) +
  scale_x_continuous(name = "Influence of variable on factor score") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        text = element_text(size = 18))

here("exhibition",
     "factors2.pdf") |>
  ggsave(big_plot,
         height = 24,
         width = 12,
         units = "in",
         device = cairo_pdf)
  
```

## Indices from factors

Figures \@ref(fig:drive-map) through \@ref(fig:amenities-map) show the spatial
variation in the drivability, walkability, density, diversity, and 
amenity-richness indices, respectively. 

```{r drive-map, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in drivability index'}
here("04_figures",
     "drivable.png") %>%
knitr::include_graphics()
```

As illistrated in Figure 
\@ref(fig:drive-map), there is an area in the center of the
region (downtown Pittsburgh) with particularly low 
drivablity (likely due to traffic congestion), surrounded 
by a ring around the center with particularly high 
drivability.

```{r walk-map, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in walkability index'}
here("04_figures",
     "walkable.png") %>%
knitr::include_graphics()
```
Figure 
\@ref(fig:walk-map) shows the opposite pattern for walkabilty, 
with particularly high walkability in the center of the
region and a ring of surrounding the center with particularly low walkability.

Figure \@ref(fig:dense-map) shows pockets of low density in the 
center of the region, possibly because the density index is
driven by residental density and these are commercial centers
with relatively few residents. With the exception of those
pockets, the density is generally highest in the center of the 
county and lowest closer to the boundaries.

```{r dense-map, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in density index'}
here("04_figures",
     "dense.png") %>%
knitr::include_graphics()
```

Figure \@ref(fig:diverse-map) suggests that the greatest diversity
(with an index representing both sociodemographic and land-use
diversity) is found in areas adjacent to the center of the region.

```{r diverse-map, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in diversity index'}
here("04_figures",
     "diverse.png") %>%
knitr::include_graphics()
```

```{r amenities-map, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in amenity-richness index'}
here("04_figures",
     "amenities.png") %>%
knitr::include_graphics()
```

Finally, Figure \@ref(fig:amenities-map) suggests an amenity-poor
area at the center of the region surrounded by and amenity-rich ring, 
with remainder of the county having an amenity richness closer to the 
county average.

Figure \@ref(fig:factor-cor) illustrates the distribution of each factor and the
relationships among them. In general, there appears to be a
trade-off between drivability and walkability and between drivability and diversity. 
There also appears to be a negative association
between density and diversity. This may be because land-use diversity was measured 
as the number of unique land uses
within the smallest buffer containing at least 2,000 residents. In very dense places,
this buffer might be too small to include a large number of unique land uses.

```{r factor-cor, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in amenity-richness index'}
here("04_figures",
     "factor-cor.png") %>%
knitr::include_graphics()
```

## Factor validation through regression

```{r, include=FALSE}
months <- c("july-2021",
            "august-2021",
            "september-2021",
            "october-2021",
            "november-2021",
            "december-2021",
            "january-2022",
            "february-2022",
            "march-2022",
            "april-2022",
            "may-2022")


permits <- here("02_data",
                "pli-permit-summary",
                     "pli-permit-summary-june-2021.xlsx") %>%
  read_xlsx(sheet = 1) %>%
  select(3, 9, 10) %>%
  rename(parcel_number = 1,
         work_type = 2, 
         structure_type = 3)
  

for (i in 1:11) {
  next_permits <- here("02_data",
                       "pli-permit-summary",
                       paste0("pli-permit-summary-",
                              months[i],
                              ".xlsx")) %>%
    read_xlsx(sheet = 1) %>%
    select(3, 9, 10) %>%
    rename(parcel_number = 1,
         work_type = 2, 
         structure_type = 3)
    
    permits <- rbind(permits, next_permits)
  
}

res_permits <- permits %>%
  mutate(PARID = paste0(substr(parcel_number, 1, 4),
                        substr(parcel_number, 6, 6),
                        substr(parcel_number, 8, 12),
                        substr(parcel_number, 14, 17),
                        substr(parcel_number, 19, 20))) %>%
  filter(work_type == "NEW CONSTRUCTION" |
           work_type == "COMPLETE DEMOLITION" |
           work_type == "NEW" |
           work_type == "NEW CONSTRUCTION" |
           work_type == "REPLACEMENT" |
           work_type == "PARTIAL DEMOLITION") %>%
  filter(structure_type == "Residential" |
           structure_type == "Residential - Two-Family")

permit_sites <- unique(res_permits$PARID)

sites <- here("02_data",
              "sites.csv") %>%
  read_csv(show_col_types = FALSE)

data <- here("02_data",
                "indices.csv") %>%
  read_csv(show_col_types = FALSE) %>%
  left_join(sites) %>%
  mutate(permit = PARID %in% permit_sites) %>%
  filter(PROPERTYCITY == "PITTSBURGH") %>%
  select(f_drivable,
         f_walkable,
         f_dense,
         f_diverse,
         f_amenities,
         permit)
```

Table \@ref(tab:regress) shows the results of three alternative regression models. The first of these is a 
null model, which assumes that likelihood of the modeled outcome (in this case, a building permit for construction
or demolition) is constant across all sites. The second model
(labeled as the "full model") predicts that likelihood based
on the variation in the three indices generated by the 
factor analysis. As shown, four of the five factors are 
statistically significant predictors of the likelihood of 
a building permit at a 99.9 percent confidence level, and this model fits the data better than the null model 
based on three measures of model fit: The Akaike Information
Criterion (AIC), the Bayesian Information Criterion (BIC), and the pseudo R^2. Based on the results of the full model, the amenity-richness index is not a significant predictor 
of the likelihood of a building permit, so a reduced model
was estimated without that variable. In this reduced model,
the remaining model coefficients were unchanged and the 
model fit was essentially unchanged.

```{r, echo=FALSE}
model_null <- glm(permit ~ 1,
             family = binomial(link = "logit"),
             data = data)

model_four <- glm(permit ~ 
               f_drivable +
               f_walkable +
               f_dense +
               f_diverse, 
             family = binomial(link = "logit"),
             data = data)

model_all <- glm(permit ~ 
               f_drivable +
               f_walkable +
               f_dense +
               f_diverse +
               f_amenities,
             family = binomial(link = "logit"),
             data = data)

model_four <- glm(permit ~ 
               f_drivable +
               f_walkable +
               f_dense +
               f_diverse, 
             family = binomial(link = "logit"),
             data = data)

export_summs(model_null, 
             model_all, 
             model_four, 
             model.names = c("Null model", "Full model", "Reduced model"),
             error_pos = "below",
             error_format = "(SE = {std.error})",
             coefs = c("(Intercept)" = "(Intercept)",
                       "Drivable" = "f_drivable",
                       "Walkable" = "f_walkable",
                       "Dense" = "f_dense",
                       "Diverse" = "f_diverse",
                       "Amenities" = "f_amenities")) %>% 
  set_caption("Results of logistic regression predicting likelihood of a demolition or construction permit over a one-year period.") %>% 
  set_label("tab:regress")
  

write_csv(tibble(variable = names(model_four$coefficients), 
                 value = model_four$coefficients),
          file = here("02_data", "model_coefs.csv"))

```

Figure \@ref(fig:effect-plot) shows how the predicted probability
of development changes as each index varies from -5 to 5 
(i.e. from five standard deviations below the average to 
five standard deviations above the average).

```{r effect-plot, echo=FALSE, warning=FALSE, fig.asp=1.0, out.width='100%', fig.cap='Effect of drivability, walkabilty, diversity, and density on development probability'}

blank_graph <- ggplot(loadings) +
  geom_point(aes(x = f_drivable,
                 y = f_walkable),
             color = NA,
             fill = NA) +
  scale_y_continuous(name = "Probability of a building permit for new construction or\ndemolition between June 2021 and May 2022") +
  theme_void() +
  theme(axis.title.y = element_text(angle = 90,
                                    face = "bold"))

effect_drive <- effect_plot(model_four, pred = f_drivable, interval = TRUE, plot.points = FALSE) +
  scale_x_continuous(limits = c(-5, 5),
                     name = "Drivable index") +
  scale_y_continuous(name = "",
                     limits = c(0, 0.1),
                     breaks = c(0, 0.02, 0.04, 0.06, 0.08, 0.1),
                     labels = c("0", "2%", "4%", "6%", "8%", "10%"))

effect_walk <- effect_plot(model_four, pred = f_walkable, interval = TRUE, plot.points = FALSE) +
  scale_x_continuous(limits = c(-5, 5),
                     name = "Walkable index") +
  scale_y_continuous(name = "",
                     limits = c(0, 0.1),
                     breaks = c(0, 0.02, 0.04, 0.06, 0.08, 0.1),
                     labels = c("0", "2%", "4%", "6%", "8%", "10%"))

effect_dense <- effect_plot(model_four, pred = f_dense, interval = TRUE, plot.points = FALSE) +
  scale_x_continuous(limits = c(-5, 5),
                     name = "Dense index") +
  scale_y_continuous(name = "",
                     limits = c(0, 0.1),
                     breaks = c(0, 0.02, 0.04, 0.06, 0.08, 0.1),
                     labels = c("0", "2%", "4%", "6%", "8%", "10%"))

effect_diverse <- effect_plot(model_four, pred = f_diverse, interval = TRUE, plot.points = FALSE) +
  scale_x_continuous(limits = c(-5, 5),
                     name = "Diverse index") +
  scale_y_continuous(name = "",
                     limits = c(0, 0.1),
                     breaks = c(0, 0.02, 0.04, 0.06, 0.08, 0.1),
                     labels = c("0", "2%", "4%", "6%", "8%", "10%"))

ggdraw() +
  draw_plot(blank_graph, x = 0, y = 0, height = 1, width = 1) +
  draw_plot(effect_drive,
            x = 0.05, y = 0.5, height = 0.5, width = 0.5) +
  draw_plot(effect_walk,
            x = 0.5, y = 0.5, height = 0.5, width = 0.5) +
  draw_plot(effect_dense,
            x = 0.05, y = 0, height = 0.5, width = 0.5) +
  draw_plot(effect_diverse,
            x = 0.5, y = 0, height = 0.5, width = 0.5) 

```

## Combined index

Recent development activity has been consistent with a hypothesis that four of the 
five indices developed from the factor analysis results represent dimensions of 
urban quality that matter to developers and property owners. We can combine these
four indices into a combined index by calculated a weighted average, where weights 
are derived from the coefficients of the regression model described above. The 
distribution of the resulting index is shown in Figure \@ref(fig:combined-hist). 
Figure \@ref(fig:combined-pitt) shows the spatial distribution of this index
across Pittsburgh, with the locations of the building permits used to estimate 
the model shown for reference.

```{r combined-hist, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Distribution of combined index, weighted according to coefficients from regression predicting the likelihood of a building permit for construction or demolition'}
here("04_figures",
     "combined-hist.png") %>%
knitr::include_graphics()
```


```{r combined-pitt, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Locations of building permits for new construction and demolition and their estimated likelihood'}
here("04_figures",
     "combined-pitt.png") %>%
knitr::include_graphics()
```

Figure \@ref(fig:combined-map) shows the spatial distribution of the combined index 
across the entire study area. It is noteworthy that there is less variation within 
the inset area than there is for any of the four indices it comprises. This is because
low scores on one index generally compensate for high scores on another.

```{r combined-map, echo = FALSE, fig.asp=1.0, out.width='100%', fig.cap='Spatial variation in combined index, weighted according to coefficients from regression predicting the likelihood of a building permit for construction or demolition'}
here("04_figures",
     "combined.png") %>%
knitr::include_graphics()
```
